---
title: "Intro to R"
author: "Allison Davis"
date: "November 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro {.tabset .tabset-fade .tabset-pills}
  
In this intro, you'll learn the basic coding skills needed to run analyses on our fish data. This is certinaly not an exhaustive list of basic codes, just the ones you'll need for the tests we will run. ***If you ever get stuck,*** google it! There are tons of resources that walk you through each step.
  
### Orientation

In R studio, your screen will be separated into 4 windows. 
  
The top right window will display 4 different tabs: environment, history, connections, and Git.

  * environment: will show you all of the current values and functions that you've created and run.
  * history: will show you all of the code actually run since you've opened R up.
  * connections: never really use, can ignore
  * git: shows you which files have been saved, and allows you to connect to your internet repository.

The bottom right window will display 5 tabs: files, plots, packages, help, viewer.

  * files: will show you all the files included in your R project folder.
  * plots: will diplay any plots run using the basic plot functions in R.
  * packages: will show you all the packages you've downloaded.
  * help: allows you to search function or package names to provide documentation (i.e. what arguments a function requires)
  * viewer: not really sure what this is to be honest...
  
The top left window is your script. Here you can type away, both functions/things you want to run as well as the annotation for each step (your descriptions on what/why you're doing a bit of code).

  * each line is associated with a number, which can help when errors say "line 49 blah blah blah"
  * if you have more than one R script open, or other document types (i.e. Rmd), these will show up as additional tabs on the top of this section
  * you can run your code by placing the cursor at the beginning or end of the line of code and hitting CTRL R
      - this is specifically a windows thing (lab laptops are windows), not sure the shortcut for Macs
      
The bottom left of your screen is your console. This is where the results of your R code will pop up. You can also directly type code into the console following the > , but this code will not be saved once you exit R -only the code saved in your R script/ R markdowns (Rmd) will be accessable.

  * there is also a terminal tab in this window. This is only useful to us when we're connecting our files to Git (online repositories), which we'll talk more about later.




### How to code

For how to install R and Rstudio, some basics on R language, and how to install pacakges try the link here: http://www.sthda.com/english/wiki/r-basics-quick-and-easy

  * this will provide some practice on the basics as well, so try the challenges!
  * another good place to review the basics: https://www.guru99.com/r-tutorial.html

Version control is important--it allows us to work on our projects on one computer, save a copy on an internet server and contine working on our project from another computer. This makes sure that if anything happens to your computer your work isn't lost! It also allows you to share your progress without having to push around files via email. You will use github and git for version control. Link 1 will lead you to github to create and account; links 2 and 3 are instructions on how to download git, and connect your Rstudio to your github.
  
  1) https://github.com/
  
  2) https://support.rstudio.com/hc/en-us/articles/200532077-Version-Control-with-Git-and-SVN
  
  3) https://happygitwithr.com/rstudio-git-github.html

When you are performing analyses that you want to share it's best to use R markdown documents. These documents allow you to present your code to people who don't have R downloaded on their computer. Instead, your code can be shared as an html, pdf, powerpoint etc. ***For this lab,*** use R markdowns (Rmd) to perform and document your analyses. For details on Rmd formating click <a href= "https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf"> here.</a>

**To review:**

It is easier to manipulate objects in R when they're assigned to a particular name.

  * to assign, you use:  <- 
      - assignments work for singular numbers, vectors of numbers, funtions, vectors of words, tables etc.
      
  * don't use an equals sign! While this will assign the object to the name correctly, it will get super confusing when you start using functions.
  
  * to view the assigned values, either type the name, or place an extra set of parentheses around the assignment code (this automatically prints the values)
  
  * Example:
```{r}
a <- 789
a

v <- c(1,3,5,7)
v

(w <- 17*3)

(farm <- c("cow", "pig","chicken"))

```

  * side note: the concatenate function [c()] allows you to combine objects (numbers, words, combos) into a single vector. If you tried to assign just the numbers 1,3,5,7 to v, R will throw an error because it does not understand that this is one group, it only sees numbers with no instructions on what to do with them. By adding c(1,3,5,7), you're telling R that you want to concatenate (or link together) these numbers, and assing that grouping or vector of numbers to the letter v.
  * If you want to replace an object with new data but the same name, just use the same name when assigning new information.

```{r}
p <- (3*2*1)
p

p <- (4*3*2*1)
p

#once you rename, the p will always be the last value!

```
  
  

### Data collection

Here I'll show you how to import the excel data from your github and into R for analyses. 

packages needed: curl

`library(curl)`

`name <- curl("url_of_your_RAW_data")`
  
  * Make sure you click the RAW button on github before copying the url --> will not work without this!

`name <- read.csv(name, header = TRUE, sep = ",", stringsAsFactors = TRUE)`
  
  * make sure the name is the same! (in both places)

`head(name)`
  
  * this will give you the first couple of rows of all columns to make sure you imported the data correctly
  
If you want to create another data frame with only certain elements, you can select these from the df you just imported in --> for example, if you wanted to separate amazons and sailfins into two different dataframes (useful for analyses on each species separately):

`amazons_raw <- name[name$X.U.FEFF.SPP =="P.form",]`

  * How to read this: from the dataframe NAME we will select[from this dataframe ($=the columns)species (== that exactly equal) "P.form", all rows]
    + by leaving the comma at the end, you ensuring that all rows that match the column specification are selected
  



### Analyses {.tabset .tabset-fade .tabset-pills}

*Select the tab that corresponds to the experiment you will analyze.*

#### Mirror experiment

Here I'll introduce some of the basic functions that we use in analyses, how to use them, and what their results tell you. 

  * summary()
  * apply() for variance and sd (mention that you can use this for any function)
  * var()
  * t.test()
  * var.test()
  * Outlier() --> need DescTools package
  * ANOVA
  * ANCOVA
  * mixed linear models?
  * linear regression?
  
#### Tit-for-tat

Here I'll introduce some of the basic functions that we use in analyses, how to use them, and what their results tell you.

  * function(x) {}
    + this is the broades type, as it allows you to create whatever function you want (particularly ones not included in basic R)
    + when creating a function, it does not matter what letters you use in the parentheses or brackets, but you must be consistent. The parentheses have the names of the variables (ex. x, y, z) and the brackets have the actual function being performed (ex. {x/2 * 100})
    + always name your function! It's useless if you can't call it back with a name.
    + example: saw we have temperature data for cities around Texas all in fahrenheit, and would like to present it in celcius
  `fahrenheit_to_celsius <- function(temp_F) {`
  `temp_C <- (temp_F - 32) * 5 / 9`
  `return(temp_C)`
`}`
        - can also be written more generally as:
  `fahrenheit_to_celsius <- function(x) {(x-32)* (5/9)}`

  * apply()
    + this allows you to apply a function to a broad amount of data all at once
    + it takes three arguments: apply(dataframe, 1OR2, function)
        - dataframe: the name of your dataframe
        - 1OR2: 1 applies the function by row; 2 applies the function by column
        - funtion: any function in basic package or that you specify (such as sum or mean)
    + example: say we have a dataframe of 30 people (rando) and 3 columns for age, weight and height.
    `apply(rando, 2, mean)`
        - this will give the average of all the ages, all the weights, and all the heights for our whole set.
      
  * lapply()
    + this is very similar to apply, but is specific for lists (will take lists and give lists). This is useful when we want to select certain columns from our dataframe.
    + it takes two arguments: lapply(list, function)
      - list: this can be a simple list that we've created or a selection of columns from our original dataframe (dataframe[2:3])
      - function: any function in basic package or one you specify
    + example: say (for whatever reason) weight and height were entered as percentages and we wanted to get the original value, we can create our own function for this.
    `lapply(rando[2:3], funtion(x) {x/100})`
      - this will divide the values for columns two and three (so weight and height) by 100.
      
  * cbind() OR rbind()
    + this stands for column bind or row bind. It allows you to add the results of a funtion or command to the end of your current dataset. 
    + it takes two arguments: cbind(original dataset name, new column name) OR rbind(original dataset name, new row name)
    + make sure you have the same number of rows or columns as the original dataset, it will report an error or place values in the incorrect spot if now!
    + example: say you wanted to add the results from you lapply function to your dataset as new columns
    `new <- lapply(rando[2:3], function(x) {x/100})`
    `rando <- cbind(rando, new_values)`
      - side note: since it is more than one column, it may not rename them using new_values. Instead it will repeat the original column names.

  * rowSums()
    + this allows you to sum all the values within a row throughout a dataset
    + it takes one argument: rowSums(name of dataset)
    + example: say you have a dataset of 10 individuals and the pieces of gum they chewed over a two week period. You want to find out how many pieces total they chewed in the second week.
    `rowSums(gum[9:15])`
      - this will give you the total number of gum pieces each individual chewed in the second week (the first column is names, 2-8 are days 1-7, so 9-15 are days 8-14). 
  
  * wilcox.test() *may not be using this, will come back to later*
  
#### Complex Maze

Here I'll introduce some of the basic functions that we use in analyses, how to use them, and what their results tell you.

  * function(x) {}
    + this is the broades type, as it allows you to create whatever function you want (particularly ones not included in basic R)
    + when creating a function, it does not matter what letters you use in the parentheses or brackets, but you must be consistent. The parentheses have the names of the variables (ex. x, y, z) and the brackets have the actual function being performed (ex. {x/2 * 100})
    + always name your function! It's useless if you can't call it back with a name.
    + example: saw we have temperature data for cities around Texas all in fahrenheit, and would like to present it in celcius
  `fahrenheit_to_celsius <- function(temp_F) {`
  `temp_C <- (temp_F - 32) * 5 / 9`
  `return(temp_C)`
`}`
        - can also be written more generally as:
  `fahrenheit_to_celsius <- function(x) {(x-32)* (5/9)}`

  * lapply()
    + this is very similar to apply, but is specific for lists (will take lists and give lists). This is useful when we want to select certain columns from our dataframe.
    + it takes two arguments: lapply(list, function)
      - list: this can be a simple list that we've created or a selection of columns from our original dataframe (dataframe[2:3])
      - function: any function in basic package or one you specify
    + example: say (for whatever reason) weight and height were entered as percentages and we wanted to get the original value, we can create our own function for this.
    `lapply(rando[2:3], funtion(x) {x/100})`
      - this will divide the values for columns two and three (so weight and height) by 100.
      
  * cbind() OR rbind()
    + this stands for column bind or row bind. It allows you to add the results of a funtion or command to the end of your current dataset. 
    + it takes two arguments: cbind(original dataset name, new column name) OR rbind(original dataset name, new row name)
    + make sure you have the same number of rows or columns as the original dataset, it will report an error or place values in the incorrect spot if now!
    + example: say you wanted to add the results from you lapply function to your dataset as new columns
    `new <- lapply(rando[2:3], function(x) {x/100})`
    `rando <- cbind(rando, new_values)`
      - side note: since it is more than one column, it may not rename them using new_values. Instead it will repeat the original column names.
  
  * t.test()
    + a t-test is used to compare the means of two groups.
    + it takes a number of arguments, but the main ones are: t.test(data$y1, data$y2, alternative = "two.sided" OR "greater" OR "less")
    + example: say you want to compare the height of men and women in the US
    `t.test(rando$hei_men, rando$hei_wom, alternative = "two.sided")`
      - we chose two.sided when the data can technically go in either direction; one sided test (greater or less) are used when it's not possible for the data to go in one direction in our comparison.
  
  * var.test()
    + the varience or F-test will compare the variation of two groups.
    + it takes a number of arguments, but like the t-test, we'll use: var.test(data$y1, data$y2, alternative = "two.sided" OR "greater" OR "less")
    + example: now we want to look at the variation of men and women's height
    `var.test(rando$hei_men, rando$hei_wom, alternative = "two.sided")`
  
  * order()
    + this will ensure categorical data, such as trials, is seen as ordered or related in R.
    `rando$trials <- ordered(rando$trials)`
  
  * lmer()
    + this function is used for mixed linear effects models on datasets that follow a normal distribution.
    + it takes three main arguments, the model, the random effect, and data: lmer(y~x1*x2 + (1|randomeffect), data=rando)
    + example: we want to see how oat growth of three different oat species is affected by nitrogen concentrations across different plots 
    `lmer(growth~species*nitro + (1|plot), data=plants)`
  
  * glmer()
    + this function is similar to linear mixed effect model, except for datasets that don't follow a normal distribution.
    + it takes the same arguments as lmer but with the addition of: family =""
      - for our datasets, we'll usually specify family="poisson" since the poisson distribution can take into account our small sample sizes.

#### Scototaxis

Here I'll introduce some of the basic functions that we use in analyses, how to use them, and what their results tell you.

* function(x) {}
    + this is the broades type, as it allows you to create whatever function you want (particularly ones not included in basic R)
    + when creating a function, it does not matter what letters you use in the parentheses or brackets, but you must be consistent. The parentheses have the names of the variables (ex. x, y, z) and the brackets have the actual function being performed (ex. {x/2 * 100})
    + always name your function! It's useless if you can't call it back with a name.
    + example: saw we have temperature data for cities around Texas all in fahrenheit, and would like to present it in celcius
  `fahrenheit_to_celsius <- function(temp_F) {`
  `temp_C <- (temp_F - 32) * 5 / 9`
  `return(temp_C)`
`}`
        - can also be written more generally as:
  `fahrenheit_to_celsius <- function(x) {(x-32)* (5/9)}`

  * apply()
    + this allows you to apply a function to a broad amount of data all at once
    + it takes three arguments: apply(dataframe, 1OR2, function)
        - dataframe: the name of your dataframe
        - 1OR2: 1 applies the function by row; 2 applies the function by column
        - funtion: any function in basic package or that you specify (such as sum or mean)
    + example: say we have a dataframe of 30 people (rando) and 3 columns for age, weight and height.
    `apply(rando, 2, mean)`
        - this will give the average of all the ages, all the weights, and all the heights for our whole set.
      
  * lapply()
    + this is very similar to apply, but is specific for lists (will take lists and give lists). This is useful when we want to select certain columns from our dataframe.
    + it takes two arguments: lapply(list, function)
      - list: this can be a simple list that we've created or a selection of columns from our original dataframe (dataframe[2:3])
      - function: any function in basic package or one you specify
    + example: say (for whatever reason) weight and height were entered as percentages and we wanted to get the original value, we can create our own function for this.
    `lapply(rando[2:3], funtion(x) {x/100})`
      - this will divide the values for columns two and three (so weight and height) by 100.
      
  * cbind() OR rbind()
    + this stands for column bind or row bind. It allows you to add the results of a funtion or command to the end of your current dataset. 
    + it takes two arguments: cbind(original dataset name, new column name) OR rbind(original dataset name, new row name)
    + make sure you have the same number of rows or columns as the original dataset, it will report an error or place values in the incorrect spot if now!
    + example: say you wanted to add the results from you lapply function to your dataset as new columns
    `new <- lapply(rando[2:3], function(x) {x/100})`
    `rando <- cbind(rando, new_values)`
      - side note: since it is more than one column, it may not rename them using new_values. Instead it will repeat the original column names.
      
* rowSums()
    + this allows you to sum all the values within a row throughout a dataset
    + it takes one argument: rowSums(name of dataset)
    + example: say you have a dataset of 10 individuals and the pieces of gum they chewed over a two week period. You want to find out how many pieces total they chewed in the second week.
    `rowSums(gum[9:15])`
      - this will give you the total number of gum pieces each individual chewed in the second week (the first column is names, 2-8 are days 1-7, so 9-15 are days 8-14). 

  * wilcox.test()
    + Wilcox rank sum tests, otherwise known as Mann Whitney U tests, is a nonparametric test (meaning the data are not normally distributed) comparing the *medians* of two separate groups. The comparison is made between a nominal characteristic (name of a group for example) and a measured characterics (body length for example). We do this comparison by assigning ranks to all the measured values in our dataset, so a 1 assigned to the lowest value, 2 to the next lowest etc, then sum up the ranks occuring in each of our groups to see how the summed ranks compare. If there is no difference between our groups, then their rank sums will be equal; if the rank sums are not equal, we can further test if they are *significantly* different. 
    + Good video describing Wilcox rank sum/Mann Whitney U: https://www.youtube.com/watch?v=fEobVCV2TJE
    + Good video differentiating common nonparametric tests (including Wilcox rank sum): https://www.youtube.com/watch?v=IcLSKko2tsg
    + it takes multiple arguments, but we'll work with four for now: wilcox.test(formula, data, alternative = c("direction"), mu)
        + formula: written as y~x, with y being the measured trait and x being the groups
        + data: name of the dataframe x and y come from 
        + alternative = c("direction"): specifying your predictions; direction can be "two.sided", "greater", or "less"
        + mu: mu is the value expected if both groups are equal (defaults to zero if not specified)
    + example: say you have discovered a new grasshopper and measured its body length. You want to see if this body length differs from another population of grasshoppers to properly call it a new species. Upon initial analyses we know the body lengths not normally distributed. We want to know if the body length of the new grasshopper is significantly different from the normal grasshopper, who has an average body length of 3.4cm.
    `new_species <- wilcox.test(bodyLength~species, data=grasshopper, alternative=c("two.sided"), mu=0)`
        + here, we are asking if bodyLength is affected by species, using our grasshopper data. Alternative is two.sided because the new grasshopper body length could technically be larger or smaller than the normal grasshopper length. Mu is 0, because we're testing the null hypothesis that the distributions of both groups are the same (so distribution x1 - distribution x2 =0).




### Graphing

There are many graphing tools both within the basic R package and through external packages. They allow you to not only display your data with different methods, but also manipulate the aesthetics of the graph.

**plot()** is the plotting tool included in the basic R package. See basic uses for plot <a href="https://www.datamentor.io/r-programming/plot-function/"> here.</a> 

While the plot function allows you to specify different types of plots, there are a number of plots that have thier own functions, such as histograms [hist()] and boxplots [boxplot()]. Check out the varieties and arguments needed <a href= "https://www.statmethods.net/graphs/index.html"> here.</a>

**ggplot()** is by far the most intense but most versitle plotting tool. To use this, you need to download and load the ggplot2 package [by library(ggplot2)]. What ggplot does is allow you to load in the basic data and then specify the type of manipulations you want to perform (do you want a bar chart or a scatterplot for example?). <a href="https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf"> Here</a> is a good cheat sheet for all that ggplot can do.




