---
title: "Intro to R"
author: "Allison Davis"
date: "November 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro
  
In this intro, you'll learn the basic coding skills needed to run analyses on our fish data. This is certinaly not an exhaustive list of basic codes, just the ones you'll need for the tests we will run. ***If you ever get stuck,*** google it! There are tons of resources that walk you through each step.
  
#### Orientation

In R studio, your screen will be separated into 4 windows. 
  
The top right window will display 4 different tabs: environment, history, connections, and Git.
  * environment:
  * history:
  * connections: never really use, can ignore
  * git:

The bottom right window will display 5 tabs: files, plots, packages, help, viewer.
  * files:
  * plots:
  * packages:
  * help:
  * viewer:
  
The top left window is your script. Here you can type away, both functions/things you want to run as well as the annotation for each step (your descriptions on what/why you're doing a bit of code).
  * each line is associated with a number, which can help when errors say "line 49 blah blah blah"
  * if you have more than one R script open, or other document types (i.e. Rmd), these will show up as additional tabs on the top of this section
  * you can run your code by placing the cursor at the beginning or end of the line of code and hitting CTRL R
      - this is specifically a windows thing (lab laptops are windows), not sure the shortcut for Macs
      
The bottom left of your screen is your console. This is where the results of your R code will pop up. You can also directly type code into the console following the > , but this code will not be saved once you exit R -only the code saved in your R script/ R markdowns (Rmd) will be accessable. 
  * there is also a terminal tab in this window. This is only useful to us when we're connecting our files to Git (online repositories), which we'll talk more about later.

#### How to code

For some basics on R language try the link here: https://difiore.github.io/applied-data-analysis-s2019/module-04/module-04.html

  * this will provide some practice on the basics as well, so try the challenges!
  
If you google and the result requires a new package, you can install a package using info in this link here: https://difiore.github.io/applied-data-analysis-s2019/module-03/module-03.html

this is link for git version control https://difiore.github.io/applied-data-analysis-s2019/module-05/module-05.html

In R you can have vectors, BLAH BLAH BLAH. (talk about each object, how they're different and what you use them for --> give examples)
  

It is easier to manipulate these objects when they're assigned to a particular name.

  * to assign, you use:  <- 
      - assignments work for singular numbers, vectors of numbers, funtions, vectors of words, tables etc.
      - make sure to talk about the c() fxn (what it means and why you need it)
      
  * don't use an equals sign! While this will assign the object to the name correctly, it will get super confusing when you start using functions.
  
  * to view the assigned values, either type the name, or place an extra set of parentheses around the assignment code (this automatically prints the values)
  
  * Example:
```{r}
a <- 789
a

v <- c(1,3,5,7)
v

(w <- 17*3)

(farm <- c("cow", "pig","chicken"))

```


#### Data collection

Here I'll show you how to import the excel data from your github and into R for analyses. 

packages needed: curl

library(curl)

name <- curl("url_of_your_RAW_data")
  make sure you click the RAW button on github before copying the url --> will not work without this!
  
name <- read.csv(name, header = TRUE, sep = ",", stringsAsFactors = TRUE)
  make sure the name is the same! (in both places)

head(name)
  this will give you the first couple of rows of all columns to make sure you imported the data correctly
  
If you want to create another data frame with only certain elements, you can select these from the df you just imported in --> for example, if you wanted to separate amazons and sailfins into two different dataframes (useful for analyses on each species separately):

amazons_raw <- name[name$X.U.FEFF.SPP =="P.form",]
  how to read this: from the dataframe NAME we will select[from this dataframe ($=the columns)species (== that exactly equal) "P.form", all rows]
  by leaving the comma at the end, you ensuring that all rows that match the column specification are selected
  



#### Analyses

Here I'll introduce some of the basic functions that we use in analyses, how to use them, and what their results tell you. 

summary()
apply() for variance and sd (mention that you can use this for any function)
var()
t.test()
hist()
var.test()
Outlier() --> need DescTools package
ANOVA
ANCOVA
mixed linear models?
linear regression?


#### Graphing

talk about basic plot in R --> what can you do, how it's useful
ggplot --> maybe include a detailed document at end at all the little things ggplot can do
  boxplots
  linear regression?
  ANOVA?



