---
title: "t4t-final"
author: "Allison Davis Connelly"
date: "4/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Predator Inspection Experiment {.tabset .tabset-fade .tabset-pills}

## New Data {.tabset .tabset-fade .tabset-pills}

This will be for the data I collect with Tristan--i.e. the tracking data.

The csv file I created from the current df in our jupyter notebook was waaaay too big to put anywhere. So, I took a few rows and created a testing csv to at least figure out how to do the analyses. 

```{r}

library(curl)

raw.track <- curl("https://raw.githubusercontent.com/allisondavis/t4t_wild-pop/master/testing.csv")

raw.track <- read.csv(raw.track, header = TRUE, sep = ",", stringsAsFactors = TRUE, row.names = NULL)

head(raw.track)



```























## Old Data {.tabset .tabset-fade .tabset-pills}

### Boldness data {.tabset .tabset-fade .tabset-pills}

#### Data collection {.tabset .tabset-fade .tabset-pills}

```{r}
library(curl)

raw.bld <- curl("https://raw.githubusercontent.com/allisondavis/t4t_wild-pop/master/bold-raw.csv")

raw.bld <- read.csv(raw.bld, header = TRUE, sep = ",", stringsAsFactors = TRUE)

head(raw.bld)

raw.bld$avg <- rowMeans(raw.bld[, c(3,4)], na.rm = TRUE)

form.bld <- raw.bld[raw.bld$SPP =="form",]
head(form.bld)

lat.bld <- raw.bld[raw.bld$SPP =="lat",]
head(lat.bld)

raw.bld.lin <- curl("https://raw.githubusercontent.com/allisondavis/t4t_wild-pop/master/bold-linear-format.csv")

raw.bld.lin <- read.csv(raw.bld.lin, header = TRUE, sep = ",", stringsAsFactors = TRUE)

form.bld.lin <- raw.bld.lin[raw.bld.lin$SPP =="form",]


lat.bld.lin <- raw.bld.lin[raw.bld.lin$SPP =="lat",]



```

*First look*

```{r}
library(rstatix)

get_summary_stats(form.bld, show = c("mean", "sd", "median", "iqr"))

get_summary_stats(lat.bld, show = c("mean", "sd", "median", "iqr"))

par(mfrow=c(1,2))

boxplot(data=raw.bld, BLD1 ~ SPP, ylab = "Exit time (s)", main= "Boldness: Round 1")

boxplot(data=raw.bld, BLD2 ~ SPP, ylab = "Exit time (s)", main= "Boldness: Round 2")


```


#### Repeatability {.tabset .tabset-fade .tabset-pills}

Here I will see how repeatable their behavior was from round 1 to round 2

```{r}
library(ggpubr)

cor.test(form.bld$BLD1, form.bld$BLD2, alternative = "two.sided", method = "spearman", use = "complete.obs")

cor.test(lat.bld$BLD1, lat.bld$BLD2, alternative = "two.sided", method = "spearman", use = "complete.obs")

form.cor <- ggscatter(form.bld, x = "BLD1", y = "BLD2", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "spearman", xlab = "Round 1", ylab = "Round 2", main = "Amazon molly")

lat.cor <- ggscatter(lat.bld, x = "BLD1", y = "BLD2", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "spearman", xlab = "Round 1", ylab = "Round 2", main = "Sailfin molly")


par(mfrow=c(1,2))
form.cor
lat.cor


```

***Amazons seem to lack repeatability (p=0.74), but sailfin have significant repeatability (p=0.0025). Not sure how to interpret this.***


Let's see if this matches with the t-test. Both are not significant-- ***this means that while they were not repeatable, the difference between the rounds was not significantly different. The averages aren't different between rounds, but the overall distribution is different for amazons (see boxplot from data collection)***

```{r}

t.test(form.bld$BLD1, form.bld$BLD2, paired = TRUE, alternative = "two.sided")

t.test(lat.bld$BLD1, lat.bld$BLD2, paired = TRUE, alternative = "two.sided")

```






#### Species comparisons {.tabset .tabset-fade .tabset-pills}

Here I will see if Amazons and sailfin differ in their boldness in round 1 and round 2. I will perform both rounds separately, but then calculate an average duration and compare. 


```{r}

t.test(form.bld$BLD1, lat.bld$BLD1, alternative = "two.sided") #sig

t.test(form.bld$BLD2, lat.bld$BLD2, alternative = "two.sided") #not sig

t.test(form.bld$avg, lat.bld$avg, alternative = "two.sided") #sig



```

***So it looks like there is a significant difference in exit time between amazons and sailfins; averages show that amazons took less time to exit than sailfins. This could influence cooperation.***


### Neophobia data {.tabset .tabset-fade .tabset-pills}

#### Data collection {.tabset .tabset-fade .tabset-pills}


```{r}
library(curl)

raw.neo <- curl("https://raw.githubusercontent.com/allisondavis/t4t_wild-pop/master/old-neo.csv")

raw.neo <- read.csv(raw.neo, header = TRUE, sep = ",", stringsAsFactors = TRUE)

head(raw.neo)

form.neo <- raw.neo[raw.neo$Spp =="form",]
head(form.neo)

lat.neo <- raw.neo[raw.neo$Spp =="lat",]
head(lat.neo)


```

*First look*

```{r}
library(rstatix)

get_summary_stats(form.neo, show = c("mean", "sd", "median", "iqr"))

get_summary_stats(lat.neo, show = c("mean", "sd", "median", "iqr"))


boxplot(data=raw.neo, Total ~ Spp, ylab = "Time to approach (s)", main="Old Object Neophobia")



t.test(form.neo$Total, lat.neo$Total, paired = FALSE, alternative = "two.sided")
var.test(form.neo$Total, lat.neo$Total, alternative = "two.sided", conf.level = 0.95) #NO DIFFERENCE IN VARIANCE p=0.374

library(ggplot2)
library(wesanderson)
#sail: #78B7C5
#amz: #E1AF00
#dark: #29211F
#med:  "#8D8680"

NEO.plt <- ggplot(raw.neo, aes(x=Spp, y=Total, fill=Spp)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) + theme(legend.title=element_blank()) + ylab("Time to approach (s)") + xlab("")
NEO.plt


```







### Cooperation Data {.tabset .tabset-fade .tabset-pills}

#### Data collection{.tabset .tabset-fade .tabset-pills} 

```{r}
library(curl)

raw.coop <- curl("https://raw.githubusercontent.com/allisondavis/t4t_wild-pop/master/pred-inspection_raw.csv")

raw.coop <- read.csv(raw.coop, header = TRUE, sep = ",", stringsAsFactors = TRUE)

head(raw.coop)

form.coop <- raw.coop[raw.coop$SPP =="form",]
head(form.coop)

lat.coop <- raw.coop[raw.coop$SPP =="lat",]
head(lat.coop)

```



*First look*

```{r}
library(rstatix)

get_summary_stats(form.coop, show = c("mean", "sd", "median", "iqr"))

get_summary_stats(lat.coop, show = c("mean", "sd", "median", "iqr"))

par(mfrow=c(1,2))

boxplot(data=raw.coop, P.INSP ~ SPP, main= "Proportion of time in inspection zone")

boxplot(data=raw.coop, P.COOP ~ SPP, main= "Proportion of time in cooperative inspection")

boxplot(data=raw.coop, P.SOLO ~ SPP, main= "Proportion of time in solo inspection")

boxplot(data=raw.coop, DEF ~ SPP, main= "Number of Defects")

```





#### Species comparisons {.tabset .tabset-fade .tabset-pills}

First, the proportion of time in the inspection zone between amazons and sailfins. 

```{r}

wilcox.test(form.coop$P.INSP, lat.coop$P.INSP, alternative = "two.sided") #not sig

wilcox.test(form.coop$P.COOP, lat.coop$P.COOP, alternative = "two.sided") #sig

wilcox.test(form.coop$P.SOLO, lat.coop$P.SOLO, alternative = "two.sided") #sig


t.test(form.coop$DEF, lat.coop$DEF, alternative = "two.sided") #not sig




```








### Pretty plots {.tabset .tabset-fade .tabset-pills}
 
To use for SICB talk...


```{r}
library(ggplot2)
library(wesanderson)
zis_colors <- wes_palette("Zissou1", type="discrete")
zis_colors2 <- wes_palette("Moonrise2", type="discrete")
zis_colors3 <- wes_palette("IsleofDogs1", type="discrete")
#sail: #78B7C5
#amz: #E1AF00
#dark: #29211F
#med:  "#8D8680"

#inspection zone
IZ <- ggplot(raw.coop, aes(x=SPP, y=P.INSP, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_violin(trim = FALSE) + stat_summary(fun.data = "mean_sdl", fun.args=list(mult=1), geom = "pointrange", color="#29211F") +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")

IZ

IN <- ggplot(raw.coop, aes(x=SPP, y=P.INSP, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")

IN


#cooperation
CP <- ggplot(raw.coop, aes(x=SPP, y=P.COOP, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_violin(trim = FALSE) + stat_summary(fun.data = "mean_sdl", fun.args=list(mult=1), geom = "pointrange", color="#29211F") +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
CP


CP1 <- ggplot(raw.coop, aes(x=SPP, y=P.COOP, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
CP1

#solo
SO <- ggplot(raw.coop, aes(x=SPP, y=P.SOLO, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_violin(trim = FALSE) + stat_summary(fun.data = "mean_sdl", fun.args=list(mult=1), geom = "pointrange", color="#29211F") +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
SO


SO1 <- ggplot(raw.coop, aes(x=SPP, y=P.SOLO, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
SO1


#boldness-R1
BD1 <- ggplot(raw.bld, aes(x=SPP, y=BLD1, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_violin(trim = TRUE) + stat_summary(fun.data = "mean_sdl", fun.args=list(mult=1), geom = "pointrange", color="#29211F") +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
BD1


BD1.1 <- ggplot(raw.bld, aes(x=SPP, y=BLD1, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
BD1.1


#boldness-R2
BD2 <- ggplot(raw.bld, aes(x=SPP, y=BLD2, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_violin(trim = FALSE) + stat_summary(fun.data = "mean_sdl", fun.args=list(mult=1), geom = "pointrange", color="#29211F") +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
BD2


BD2.1 <- ggplot(raw.bld, aes(x=SPP, y=BLD2, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
BD2.1


#boldness-avg
AV <- ggplot(raw.bld, aes(x=SPP, y=avg, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_violin(trim = FALSE) + stat_summary(fun.data = "mean_sdl", fun.args=list(mult=1), geom = "pointrange", color="#29211F") +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
AV


AV1 <- ggplot(raw.bld, aes(x=SPP, y=avg, fill=SPP)) + scale_fill_manual(values = c( "#E1AF00", "#78B7C5")) + geom_boxplot(outlier.color = "#29211F", outlier.shape = 16, outlier.size = 2, notch = FALSE) + geom_point(color= "#8D8680", size=2, alpha=0.5) +theme_classic()  + scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) + theme(legend.title=element_blank()) + ylab("") + xlab("")
AV1



ggsave(IN, filename = "Inspection.png", bg="transparent", width = 180, height = 120, units = "mm", dpi = 300)
ggsave(CP1, filename = "Cooperation.png", bg="transparent", width = 180, height = 120, units = "mm", dpi = 300)
ggsave(SO1, filename = "Solo.png", bg="transparent", width = 180, height = 120, units = "mm", dpi = 300)
ggsave(BD1.1, filename = "BoldR1.png", bg="transparent", width = 180, height = 120, units = "mm", dpi = 300)
ggsave(BD2.1, filename = "BoldR2.png", bg="transparent", width = 180, height = 120, units = "mm", dpi = 300)
ggsave(AV1, filename = "Bold.Avg.png", bg="transparent", width = 180, height = 120, units = "mm", dpi = 300)






```


