---
title: "t4t-final"
author: "Allison Davis Connelly"
date: "4/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data collection 

```{r}
library(curl)

raw <- curl("https://raw.githubusercontent.com/allisondavis/t4t_wild-pop/master/t4t_quick-look_SUTS.csv")

raw <- read.csv(raw, header = TRUE, sep = ",", stringsAsFactors = TRUE)

raw <- raw[-c(15:26)]
raw <- raw[-c(11:987),]
head(raw)

form <- raw[raw$SPP =="form",]
form <- form[-c(15:26)]
head(form)

lat <- raw[raw$SPP =="lat",]
lat <- lat[-c(15:26)]
head(lat)


```

## Analyses {.tabset .tabset-fade .tabset-pills}

### Calculating the average percent of occurances in each section

```{r}

#get total possible obs number

form_sums <- rowSums(form[5:14]) 
form <- cbind(form, form_sums)

lat_sums <- rowSums(lat[5:14])
lat <- cbind(lat, lat_sums)


#get avg numb spent in each section for each indiv

form_pct.id <- lapply(form[5:14], function(x){x/form_sums}) 
form <- cbind(form, form_pct.id)

lat_pct.id <- lapply(lat[5:14], function(x) {x/lat_sums})
lat <- cbind(lat, lat_pct.id)
lat

#get avg numb spent in each section for whole spp and covert to percent

form_avg <- apply(form[16:25], 2, mean) 
form_pct.spp <- form_avg*100                

lat_avg <- apply(lat[16:25], 2, mean)
lat_pct.spp <- lat_avg*100


#isolate just the percentages for each section to plot

pcts.spp <- rbind(form_pct.spp, lat_pct.spp)
pcts.spp


#to plot avg times spent in each section per species
x <- c(1,2,3,4,5,6,7,8,9,10)

df <- data.frame(x, form_pct.spp, lat_pct.spp)

library(ggplot2)

ggplot(df, aes(x)) + geom_line(aes(y=form_pct.spp, color="Amazon"), size= 1.5) + geom_line(aes(y=lat_pct.spp, color="Sailfin"), size = 1.5)  +ylab("Times in section (%)\n") + xlab("Sections") + geom_point(aes(y=form_pct.spp, color="Amazon"),size=2, shape=21, fill="white") + geom_point(aes(y=lat_pct.spp, color="Sailfin"), size=2, shape=21, fill="white") + scale_shape(solid=FALSE) +theme_classic() + scale_color_manual(values = c("#9C964A", "#FAD77B")) + scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) + theme(legend.title=element_blank()) + theme(text = element_text(size = 16))


```


### Calculating frequency of occurances in 3 sections: safety, neutral, danger
  
   - would I have to sum the proportions instead? Since they don't all have the same total number of observations?


```{r}
#sum occurances for sections 1-3(danger), 4-6(neutral), and 7-10(safety)

raw.dan.num <- rowSums(raw[5:7])
raw <- cbind(raw, raw.dan.num)

raw.neu.num <- rowSums(raw[8:10])
raw <- cbind(raw, raw.neu.num)

raw.saf.num <- rowSums(raw[11:14])
raw <- cbind(raw, raw.saf.num)

#wilcox rank sum tests (aka mann whitney u tests) to see if there's a significant difference between species within each section
raw.dan <- wilcox.test(raw.dan.num~raw$SPP)
raw.dan

raw.neu <- wilcox.test(raw.neu.num~raw$SPP)
raw.neu

raw.saf <- wilcox.test(raw.saf.num~raw$SPP)
raw.saf


```

* <font color = "darkorchid4"> There does not appear to be a significant difference in the occurrences within three sections (danger, neutral, safety) between the two species.</font>
    + not sure if this was the correct way to calculate, since the occurrences are not independent -if you are in the neutral, you had to pass through the safety section to get there. 

<br><br><br>
* Going to repeat this analysis but with the proportions instead


```{r}

#get percents for each section based on total observation number

sums <- rowSums(raw[5:14]) 
raw <- cbind(raw, sums)

pct <- lapply(raw[5:14], function(x){x/sums}) 
raw <- cbind(raw, pct)


#sum occurances for sections 1-3(danger), 4-6(neutral), and 7-10(safety)

raw.dan.num <- rowSums(raw[16:18])
raw <- cbind(raw, raw.dan.num)

raw.neu.num <- rowSums(raw[19:21])
raw <- cbind(raw, raw.neu.num)

raw.saf.num <- rowSums(raw[22:25])
raw <- cbind(raw, raw.saf.num)

#wilcox rank sum tests (aka mann whitney u tests) to see if there's a significant difference between species within each section
raw.dan <- wilcox.test(raw.dan.num~raw$SPP)
raw.dan

raw.neu <- wilcox.test(raw.neu.num~raw$SPP)
raw.neu

raw.saf <- wilcox.test(raw.saf.num~raw$SPP)
raw.saf



```


Seems to at least not throw an error as compared to the regular scores. Still no significant difference. 
